---
title: "Concentration estimation"
output: html_document
---

```{r}
library(reticulate)
library(data.table)
library(ggplot2)
py_require(packages = c("numpyro",
                        "blackjax",
                        "IPython"))
```


# Prep the data
Let's get the data for Chamber of Deputees.
For the data we need to get shares for each of the relevant parties:
  Movimento 5 Stelle,
  Partito Democratico,
  Lega,
  Fratelli d'Italia,
  Forza Italia,
  Others,
at electoral district (COMUNE) level.

```{r}
source("../explore_data/utils.R")
#load the data from eligendo.it
load_all_data_tables()

#perform some cleaning for the names
set_region_column_camera()
set_region_column_senato()
```

Get the share of each party of interest in each electoral station and abstension (1 - turnout). Abstension will be modelled as a party. 
It is relevant as Italian elections experienced a strong decrease in turnout in recent elections.
```{r}
data22 <- party_abs_vote_share_by_comune(camera22)
data18 <- party_abs_vote_share_by_comune(camera18)
data13 <- party_abs_vote_share_by_comune(camera13)

data22
```

Let's visualize our data:

The Democratic Party is the historical center-left party. We can compare the distribution of polling share distribution in two region: EMILIA, stronghold of the left; VENETO, stronghold of the right.
```{r}
data22[REGION == "EMILIA" | REGION == "VENETO"] |> 
  ggplot(aes(x = PARTITO_DEMOCRATICO))+
  geom_density(alpha = 0.7, fill = "red3")+
  geom_freqpoly(aes(y = after_stat(density)), binwidth = 0.005)+
  theme_bw()+
  theme(legend.position = "none")+
  facet_grid(REGION~.)+
  xlim(c(0,1))
```

We can do the same thing for the right wing party Fratelli d'Italia.
```{r}
data22[REGION == "EMILIA" | REGION == "VENETO"] |> 
  ggplot(aes(x = FRATELLI_D_ITALIA))+
  geom_density(alpha = 0.7, fill = "green4")+
  geom_freqpoly(aes(y = after_stat(density)), binwidth = 0.005)+
  theme_bw()+
  theme(legend.position = "none")+
  facet_grid(REGION~.)+
  xlim(c(0,1))
```

Or even for Abstension:
```{r}
data22[REGION == "EMILIA" | REGION == "VENETO"] |> 
  ggplot(aes(x = ABSTENTION))+
  geom_density(alpha = 0.7, fill = "brown2")+
  geom_freqpoly(aes(y = after_stat(density)), binwidth = 0.005)+
  theme_bw()+
  theme(legend.position = "none")+
  facet_grid(REGION~.)+
  xlim(c(0,1))
```


# Now let's find a way to estimate the data by region.
```{python}
from jax import random, vmap
import jax.numpy as jnp
from jax.scipy.special import logsumexp

import numpyro
from numpyro import handlers
from numpyro.diagnostics import hpdi
import numpyro.distributions as dist
from numpyro.infer import MCMC, NUTS

import numpy as np


def model_region(obs = None, n_district = None):
    
    with numpyro.plate("party", 4):
      # log_a = numpyro.sample("log_a", dist.Normal(0, 3))
      a = numpyro.sample("a", dist.HalfNormal(3))
      
    # a = jnp.exp(log_a)
    
    with numpyro.plate("data", 160):
      numpyro.sample("obs", dist.Dirichlet(a), obs = obs)
        

```

```{r}
nuts_kernel = py$NUTS(py$model_region)

mcmc = py$MCMC(nuts_kernel, num_samples=2000L, num_warmup=2000L, jit_model_args = TRUE )
rng_key = py$random$key(42L)

# Get the data in jax arrays
data_jax = py$jnp$array(as.matrix(data[data$region == 3, 1:4]))
data_jax$shape[1] |> print()

mcmc$run(rng_key, obs = data_jax, n_district = py$jnp$array(160L))
```

```{r}
mcmc$print_summary()

c(-0.72, 0.03, 1.16, 1.92) |> exp()
c(0.08, 0.06, 1.65, 1.37) |> exp()
c(-0.97, -0.03, 1.73, 1.48) |> exp()
```


# Now lejt's build the hierarchical model.
```{python}
# def model_national(obs = None, region_idx = None, N = None):
#     
#     # Make the prior for the population dirichlet
#     # with numpyro.plate("party", 4):
#     #   a = numpyro.sample("a", dist.HalfNormal(1))
#       
#     with numpyro.plate("region", len(np.unique(region_idx))):
#       a_reg = numpyro.sample("a_reg", dist.Dirichlet)
#       
#     with numpyro.plate("data", N):
#       numpyro.sample("obs", dist.Dirichlet(a_reg[region_idx]), obs = obs)
#       
      
def model_national(obs = None, region_idx = None, N = None):
    
    sigma_reg = numpyro.sample("sigma_reg", dist.HalfNormal(1))
    
    with numpyro.plate("party", 4):
      log_c_party = numpyro.sample("c_party", dist.Normal(jnp.zeros(1), jnp.ones(1)*3))
    
    c_party = jnp.exp(log_c_party)
    
    with numpyro.plate("party", 4):
      with numpyro.plate("region", len(np.unique(region_idx))):
        log_a_reg = numpyro.sample("a_reg", dist.Normal(c_party, sigma_reg))
      
    a_reg = jnp.exp(log_a_reg)
    
    with numpyro.plate("data", N):
      numpyro.sample("obs", dist.Dirichlet(a_reg[region_idx]), obs = obs)
```

```{r}
nuts_kernel = py$NUTS(py$model_national)

mcmc = py$MCMC(nuts_kernel, num_samples=2000L, num_warmup=1000L, jit_model_args = TRUE)
rng_key = py$random$key(42L)

# Get the data in jax arrays
data_jax = py$jnp$array(as.matrix(data[, 1:4]))
region_idx =  py$jnp$array(as.integer(data[, 5]))
```


# Move it to blackjax
```{python}
from numpyro.infer.util import initialize_model

rng_key, init_key = random.split(r.rng_key)

init_params, potential_fn_gen, *_ = initialize_model(init_key, model_national, model_args=(r.data_jax, r.region_idx, 1526), dynamic_args=True, )
```

Convert it to a logdensity function:
```{python}
from jax import jit

logdensity_fn = lambda position: -potential_fn_gen(r.data_jax, r.region_idx, 1526)(position)
logdensity_jit = jit(logdensity_fn)
initial_position = init_params.z
```

Create the kernel for nuts algo with step size and inv mass matrix
```{python}
import blackjax

inv_mass_matrix = jnp.ones(45)
step_size = 1e-3

nuts = blackjax.nuts(logdensity_jit, step_size, inv_mass_matrix)
init_state = nuts.init(initial_position)
kernel = nuts.step

```
 the inference loop
```{python}
import jax

def inference_loop(rng_key, kernel, initial_state, num_samples):
  @jax.jit
  def one_step(state, rng_key):
      state, _ = kernel(rng_key, state)
      return state, state

  keys = jax.random.split(rng_key, num_samples)
  _, states = jax.lax.scan(one_step, initial_state, keys)

  return states
```

Run the algo:
```{python}
n_samples = 100
rng_key, sample_key = random.split(rng_key)
states = inference_loop(sample_key, kernel, init_state, n_samples)

mcmc_samples = states.position
```

Save the results of the test to file:
```{python}
# 
# import pickle
# import os
# 
# path = "C:/Users/Ermenegisto/CODING/R/UNI/CSS/final_proj/dirichlet_estimation"
# with open(path, 'wb') as file:
#     pickle.dump(mcmc_samples, file)
```



# Let's look at the results in r
```{r}
np = import("numpy", as = "np")
output = py$mcmc_samples

# population level estimates, prior
data_pop = as.data.frame(np$array(output$c_party))
names(data_pop) <- c("a1","a2","a3","a4")

library(tidyr)

long <- data_pop |> # [1000:nrow(data_pop)] %>%
  pivot_longer(
    cols = a1:a4, 
    names_to = "alpha",
    values_to = "value"
)

alpha
exp(sapply(data_pop, mean))

long |> 
  ggplot(aes(x = exp(value)))+
  geom_density()+
  facet_wrap(.~alpha)+
  theme_bw()
```

And now let's plot the region based ones
```{r}
data_reg = output$a_reg

n_samples = data_reg$shape[[1]]
burnout = 10

data_plot = data_reg[burnout:n_samples,,]
# mean_concentration per region x party
estimates = as.data.frame(np$array(data_plot$mean(0L)))
names(estimates) <- party_names
exp(estimates)
```


```{r}
actractiveness_by_region
```


```{r}
# %error (using %change/size = ln(a) - ln(b) for small %)
sapply(log(actractiveness_by_region) - estimates , mean)
```

Save the data in R format:
```{r}
r_output = lapply(output, function(jx){
  np$array(jx)
})

saveRDS(r_output, file = "sample_model_1.rds")
```




```{r}
rnorm(100, 0, 4) |> exp() |> 
  ggplot(aes(x))
```




