---
title: "Concentration estimation"
output: html_document
---

```{r}
library(reticulate)
library(data.table)
library(dplyr)
library(ggplot2)
library(DirichletReg)
library(compositions)
py_require(packages = c("numpyro",
                        "blackjax",
                        "IPython"))
```


# Prep the data
Let's get the data for Chamber of Deputees.
For the data we need to get shares for each of the relevant parties:
  Movimento 5 Stelle,
  Partito Democratico,
  Lega,
  Fratelli d'Italia,
  Forza Italia,
  Others,
at electoral district (COMUNE) level.

```{r}
source("../R/utils.R")
#load the data from eligendo.it
load_all_data_tables()

#perform some cleaning for the names
set_region_column_camera()
set_region_column_senato()

camera22[COLLEGIO == "PIEMONTE 2 - U02 (NOVARA)", .(filter_minor_parties(LISTA), VOTI_LISTA, ELETTORI)]
```

Get the share of each party of interest in each electoral station and abstension (1 - turnout). Abstension will be modelled as a party. 
It is relevant as Italian elections experienced a strong decrease in turnout in recent elections.
```{r}
data22 <- party_abs_vote_share_by_comune(camera22)
# parties = data22[,-c(1,2,3,4)] |> names()
data18 <- party_abs_vote_share_by_comune(camera18)
data13 <- party_abs_vote_share_by_comune(camera13)

data18 <- data18[-316]
data13 <- data18[-316]
# over_share22 = (data22[,-c(1,2,3,4)] |> apply(1, sum) > 1) |> which()
# over_share18 = (data18[,-c(1,2,3,4)] |> apply(1, sum) > 1) |> which()
# over_share13 = (data13[,-c(1,2,3,4)] |> apply(1, sum) > 1) |> which()

camera13[COLLEGIO == "09 - PINEROLO" & COMUNE == "SAUZE DI CESANA", .(PARTY = filter_minor_parties(LISTA), VOTI_LISTA, ELETTORI)]
```

Let's visualize our data:

The Democratic Party is the historical center-left party. We can compare the distribution of polling share distribution in two region: EMILIA, stronghold of the left; VENETO, stronghold of the right.
```{r}
data22[REGION == "EMILIA" | REGION == "VENETO"] |> 
  ggplot(aes(x = PARTITO_DEMOCRATICO))+
  geom_density(alpha = 0.7, fill = "red3")+
  geom_freqpoly(aes(y = after_stat(density)), binwidth = 0.005)+
  theme_bw()+
  theme(legend.position = "none")+
  facet_grid(REGION~.)+
  xlim(c(0,1))
```

We can do the same thing for the right wing party Fratelli d'Italia.
```{r}
data22[REGION == "EMILIA" | REGION == "VENETO"] |> 
  ggplot(aes(x = FRATELLI_D_ITALIA))+
  geom_density(alpha = 0.7, fill = "green4")+
  geom_freqpoly(aes(y = after_stat(density)), binwidth = 0.005)+
  theme_bw()+
  theme(legend.position = "none")+
  facet_grid(REGION~.)+
  xlim(c(0,1))
```

Or even for Abstension:
```{r eval=FALSE, include=FALSE}
data22[REGION == "EMILIA" | REGION == "VENETO"] |> 
  ggplot(aes(x = ABSTENTION))+
  geom_density(alpha = 0.7, fill = "brown2")+
  geom_freqpoly(aes(y = after_stat(density)), binwidth = 0.005)+
  theme_bw()+
  theme(legend.position = "none")+
  facet_grid(REGION~.)+
  xlim(c(0,1))
```


beware of observation 316.

Let's fit using the dirichletReg package

Let's try to fit the dirichlet by region first:
```{r}
gtrend = load_gtrends_data()
rcsr = get_rcsr_by_region(gtrend, year = 2018)

# Timestep is parliamentary cycle:

data13$year = 0
data18$year = 1
data22$year = 2

# Create the "treatment" overtly far right communication variable for 
# the parties "LEGA" and "FRATELLI_D_ITALIA" (LEGA before 2018 election and fratelli before 2018)

data13$treat_lega = 0
data18$treat_lega = 1
data22$treat_lega = 1

data13$treat_frat = 0
data18$treat_frat = 0
data22$treat_frat = 1

# include encoding for incumbency for all big parties
# pdl (same as forza italia) was in government with the league. Members of fi where also part of the goverment in the "super party" pdl
inc13 = data.frame(inc_pd = 0, inc_m5s = 0, inc_lega = 1, inc_fd = 1, inc_fi = 1)
# forza italia made a short lived government alliance with the center left, that's why it is included in the incumbency for 2018
inc18 = data.frame(inc_pd = 1, inc_m5s = 0, inc_lega = 0, inc_fd = 0, inc_fi = 1)
# the only party who didn't take part into the government coalition was fratelli d'italia
inc22 = data.frame(inc_pd = 1, inc_m5s = 1, inc_lega = 1, inc_fd = 0, inc_fi = 1)

data13_inc = cbind(data13, inc13)
data18_inc = cbind(data18, inc18)
data22_inc = cbind(data22, inc22)



data_inc = rbind(data13_inc, data18_inc, data22_inc)

# data_inc <- data_inc[REGION == "EMILIA",]
# data_inc <- data22_inc

data_inc |> names()

# conflict with the data.table class, change it
class(data_inc) <- class(data_inc)[2]

# change the names so that they comply with brms param names
data_inc$PD = data_inc$PARTITO_DEMOCRATICO
data_inc$M5S = data_inc$MOVIMENTO_5_STELLE
data_inc$FRI = data_inc$FRATELLI_D_ITALIA
data_inc$FI = data_inc$FORZA_ITALIA

data_inc$REGION = factor(data_inc$REGION)


tmp <- with(data_inc, cbind(FI, FRI,LEGA, M5S , O, PD, ABSTENTION))

# use the DirichletReg conversion function to make sure data fits onto the simplex.
tmp <- DR_data(tmp)
data_inc$prop = as.matrix(tmp)


```

Fit Dir using the composition library
```{r}
# library(compositions) 


# function for estimating dirichlet params by region given datasets

get_actractiveness_dt <- function(DT, ...){
  
  
  
  fitted = lapply(unique(DT$REGION), function(reg){
    param = fitDirichlet(DT[REGION == reg, .(ABSTENTION, FORZA_ITALIA,
                                             FRATELLI_D_ITALIA, LEGA,
                                             MOVIMENTO_5_STELLE, O,
                                             PARTITO_DEMOCRATICO)])
    data.frame(alpha = param$alpha, REGION = reg, PARTY = ...)
  })
  
  
  rbindlist(fitted)
}

party_names = c("ABS", "FI", "FDI", "LEGA", "M5S", "OTH", "PD")

al13 = get_actractiveness_dt(data13, PARTY = party_names, year = 0)
al18 = get_actractiveness_dt(data18, PARTY = party_names, year = 1)
al22 = get_actractiveness_dt(data22, PARTY = party_names, year = 2)

data_alpha = rbind(al13, al18, al22)

```


Now join with the auxiliary variables:
```{r}
#reduced dataset for the join
# names(data_inc)
# [1] "ELETTORI"            "COLLEGIO"            "REGION"              "COMUNE"             
#  [5] "ABSTENTION"          "FORZA_ITALIA"        "FRATELLI_D_ITALIA"   "LEGA"               
#  [9] "MOVIMENTO_5_STELLE"  "O"                   "PARTITO_DEMOCRATICO" "year"               
# [13] "treat_lega"          "treat_frat"          "inc_pd"              "inc_m5s"            
# [17] "inc_lega"            "inc_fd"              "inc_fi"              "PD"                 
# [21] "M5S"                 "FRI"                 "FI"                  "prop"  

tmp = data_alpha |> 
  mutate(M = if_else( (PARTY == "LEGA" & year > 0) | (PARTY == "FDI" & year > 1), 1, 0),
         inc = case_when(
           PARTY == "PD" & (year == 1 | year == 2) ~ 1 ,
           PARTY == "M5S" & year == 2 ~ 1 ,
           PARTY == "FI" ~ 1 ,
           PARTY == "FDI" & year == 0 ~ 1 ,
           PARTY == "PD" & (year == 0 | year == 2) ~ 1 ,
           .default = 0
         ))

data_1 = inner_join(tmp, rcsr, by = "REGION")
data_no_abs = data_1[PARTY != "ABS", ]
```


```{r}
library(lme4)
fit2= lmer(log(alpha) ~ (1 | REGION:PARTY) + year + year^2 + M + inc + RCSR_2018:M , data_no_abs, REML = FALSE)
# fitlm= lm(log(alpha) ~ REGION:PARTY + PARTY:year + M + inc + RCSR_2018:M , data_no_abs, REML = FALSE)
fit1= lmer(log(alpha) ~ (1 | REGION:PARTY) + (year + year^2 | PARTY) + M + inc  , data_no_abs, REML = FALSE)

plot(fit2)

anova(fit2, fit1)
plot(fit1, resid(.)  ~  log(alpha) | PARTY)
plot(fit1, resid(.)  ~  log(alpha) | year)
plot(fit1,PARTY   ~resid(.)  | year)
plot(fit1, fitted(.)  ~  log(alpha) | PARTY)
plot(fit1, fitted(.)  ~  log(alpha) | year)
plot(fit1, resid(., scale = TRUE)  ~  fitted(.)| M)
```


Now try the model where the "treatment variable" is the switch in comunication rather than the comunication itself.
```{r}

data_inc_2 = data_alpha |> 
  mutate(M = if_else( (PARTY == "LEGA" & year == 1) | (PARTY == "FDI" & year == 2), 1, 0),
         inc = case_when(
           PARTY == "PD" & (year == 1 | year == 2) ~ 1 ,
           PARTY == "M5S" & year == 2 ~ 1 ,
           PARTY == "FI" ~ 1 ,
           PARTY == "FDI" & year == 0 ~ 1 ,
           PARTY == "PD" & (year == 0 | year == 2) ~ 1 ,
           .default = 0
         ))

data_2 = inner_join(data_inc_2, rcsr, by = "REGION")
  
library(lme4)
fit_1= lmer(log(alpha) ~ (1 | REGION:PARTY) + (1 | PARTY:year) + M + inc + RCSR_2018:M , data_2, REML = FALSE)
fit_2= lmer(log(alpha) ~ (1 | REGION:PARTY) + (1 | PARTY:year) + M + inc  , data_2, REML = FALSE)

anova(fit_1, fit_2)
```



fit test regression on lme4
```{r}
library(lme4)

glmer(alpha ~ (1|REGION) + inc_pd + inc_m5s , data = data_no_rcsr, family = Gamma)
b


make_standata(bf(alpha~ (1 | REGION:PARTY) + (1| PARTY:year) + inc_pd + inc_m5s +
                   inc_fi + year) ,
              data_no_rcsr,
              exponential(link = "log"))
fit = brm(bf(alpha~ (1 | REGION:PARTY) + (1| PARTY:year) + inc_pd + inc_m5s +
                   inc_fi + year) ,
              data_no_rcsr,
              exponential(link = "log"))
```

```{r}
summary(fit)
```



```{r}
library(brms)

# to avoid confusion with brms internals
bind <- function(...) cbind(...)

# fit <- brm(bf(bind(FORZA_ITALIA, FRATELLI_D_ITALIA,LEGA, MOVIMENTO_5_STELLE , O, PARTITO_DEMOCRATICO, ABSTENTION) ~ year + , muFORZA_ITALIA ~ inc_fi), data_inc, dirichlet())
# make_standata(bf(prop ~ year,
#                  muFRI~ inc_fd + treat_frat, muM5S~inc_m5s, muPD~inc_pd, muLEGA~inc_lega + treat_lega, muFI~inc_fi),
#               data_inc,
#               dirichlet(refcat = "ABSTENTION"))
fit1 = brm(bf(prop ~ year,
                 muFRI~ inc_fd + treat_frat, muM5S~inc_m5s, muPD~inc_pd, muLEGA~inc_lega + treat_lega, muFI~inc_fi),
              data_inc,
              dirichlet(refcat = "ABSTENTION"))
```


#fit the dirichlet:


```{r}
lm(PARTITO_DEMOCRATICO ~ inc_pd + year + FORZA_ITALIA + LEGA + FRATELLI_D_ITALIA + REGION, data_inc) -> test_lm
test_lm |> summary()

animus = inner_join(, data13)
party = DR_data(animus[,-c(1,2,3,4,5)])

animus$test = rnorm(nrow(animus))
animus$test_2 = 2 * rnorm(nrow(animus))
animus$test_3 =   rnorm(nrow(animus))

fit_test = DirichReg(party~RCSR_2022 + test * test_2 +  test_3:RCSR_2022, animus)
tmp
```

```{r}
fit_test |> summary()
fit_test$formula
```














Import the necessary packages.
```{python}
from jax import random, vmap
import jax.numpy as jnp
from jax.scipy.special import logsumexp

import numpyro
from numpyro import handlers
from numpyro.diagnostics import hpdi
import numpyro.distributions as dist
from numpyro.infer import MCMC, NUTS, HMC

import numpy as np
```


Define the model.
```{python}
def model_region(obs = None):
    
    with numpyro.plate("party", int(obs.shape[1])):
      # log_a = numpyro.sample("log_a", dist.Normal(0, 3))
      log_a = numpyro.sample("a", dist.HalfNormal(10))
    
    a = jnp.exp(log_a)
    
    with numpyro.plate("data", int(obs.shape[0])):
      numpyro.sample("obs", dist.Dirichlet(a), obs = obs)
        

```

Initialize the Numpyro NUTS kernel.
```{r}
nuts_kernel = py$HMC(py$model_region)

mcmc = py$MCMC(nuts_kernel, num_samples=4000L, num_warmup = 4000L, num_chains = 1L, jit_model_args = TRUE,)
rng_key = py$random$key(42L)
```

Quick std for proportions (some issue in data aggregation for small polling stations) implementation:
```{r}
std_1 <- function(x){
  x/sum(x)
}
# keep proper values the same
std_1(c(0.4,0.4,0.2)) == c(.4,.4,.2)
# adjust for aggregation/numerical errors for smaller locations
std_1(c(0.5,0.5,0.2)) |> sum() == 1
```

let's fit the dirichlet:
```{r}

party_names <- data22[,-c(1,2,3,4)] |> names()

# fit the dirchlet for all regions:
fitted = lapply(unique(data22$REGION), function(reg){
  # fit the dirichlet
  param = fitDirichlet(data22[REGION == reg, -c(1,2,3,4)])$alpha
  # get the number of polling stations
  N_pol =  data22[REGION == reg, .N]
  list(param = param,
       conc = sum(param),
       REGION = reg,
       n_polling = N_pol)
})

#organize it in a tabular dataset:

alpha_by_party = as.data.frame(t(sapply(fitted, function(fit){
  fit$param
})))

# get the other variables:
rest_var = lapply(fitted, function(fit){
  list(conc = fit$conc, REGION = fit$REGION, n_polling = fit$n_polling)
})

rest_dt = rbindlist(rest_var)

names(alpha_by_party) <- party_names
data_y = cbind(alpha_by_party, rest_dt)
data_y 
```


Let's get some visualization in this data

In the following graph we can observe that the #polling station doesn't affect the concentration of the votes.
```{r}
data_y |> 
  ggplot(aes(x = n_polling, y = conc, colour = REGION))+
  geom_point()+
  theme_bw()
```

Let's see the correlation between the actractiveness of fratelli d'italia and actractiveness of abstention 
```{r}
data_y |> 
  ggplot(aes(x = ABSTENTION, y = FRATELLI_D_ITALIA, colour = REGION))+
  geom_point()+
  theme_bw()
```

Now try the same with the democratic party


```{r}
data_y |> 
  ggplot(aes(x = ABSTENTION, y = PARTITO_DEMOCRATICO, colour = REGION))+
  geom_point()+
  theme_bw()
```





Try to convert data into long format:
```{r}
long_data = melt(as.data.table(data_y), measure.var = c(1,2,3,4,5,6,7))
long_data 
```

try the racial animus variable
```{r}
gtrend = load_gtrends_data()
RCSR_2022 = get_rcsr_by_region(gtrend, year= 2022)
fin_dat = inner_join(long_data[variable == "FRATELLI_D_ITALIA",],
           RCSR_2022, by = "REGION")

lm(value~RCSR_2022,fin_dat) |> summary()
```



Convert the data to jax format.
```{r}
jnp = import("jax.numpy", as = "jnp")
# correct_matrix <- as.matrix(data22[REGION == "EMILIA",-c(1,2,3,4)])
# std_matrix = t(apply(t(correct_matrix), 2, std_1))
data_veneto_raw = jnp$array(as.matrix(data22[REGION == "VENETO",-c(1,2,3,4)]))
data_em_raw = jnp$array(as.matrix(data22[REGION == "CALABRIA",-c(1,2,3,4)]))

ve = py$np$array(data_veneto_raw)
em = py$np$array(data_em_raw)


data_emilia = data_veneto_raw
data_emilia = data_em_raw
# (data_emilia - data_em_raw)
# data_emilia = data_emilia[1:50,]
```

```{r}
# adjust the prng structure over the doc NOTE

rng_key = py$random$key(42L)


mcmc$run(rng_key, obs = data_emilia)
```

```{r}
mcmc$print_summary()
```
```{r}
data22[REGION == "VENETO",]
```

# Move it to blackjax
```{python}
from numpyro.infer.util import initialize_model

rng_key, init_key = random.split(r.rng_key)

init_params, potential_fn_gen, *_ = initialize_model(init_key, model_region, model_args=(r.data_emilia,), dynamic_args=True, )
```

Convert it to a logdensity function:
```{python}
from jax import jit

logdensity_fn = lambda position: -potential_fn_gen(r.data_emilia)(position)
logdensity_jit = jit(logdensity_fn)
initial_position = init_params.z
```

Create the kernel for nuts algo with step size and inv mass matrix
```{python}
import blackjax

inv_mass_matrix = jnp.ones(7)
step_size = 1e-3

nuts = blackjax.nuts(logdensity_jit, step_size, inv_mass_matrix)
init_state = nuts.init(initial_position)
kernel = nuts.step

```
 the inference loop
```{python}
import jax

def inference_loop(rng_key, kernel, initial_state, num_samples):
  @jax.jit
  def one_step(state, rng_key):
      state, _ = kernel(rng_key, state)
      return state, state

  keys = jax.random.split(rng_key, num_samples)
  _, states = jax.lax.scan(one_step, initial_state, keys)

  return states
```

Run the algo:
```{python}
n_samples = 10000
rng_key, sample_key = random.split(rng_key)
states = inference_loop(sample_key, kernel, init_state, n_samples)

mcmc_samples = states.position
```

Save the results of the test to file:
```{python}
# 
# import pickle
# import os
# 
# path = "C:/Users/Ermenegisto/CODING/R/UNI/CSS/final_proj/dirichlet_estimation"
# with open(path, 'wb') as file:
#     pickle.dump(mcmc_samples, file)
```




And now let's plot the region based ones
```{r}
data_reg = py$mcmc_samples$a
data_plot = as.data.frame(np$array(data_reg))

names(data_plot)

data_plot |> 
  ggplot(aes())+
  geom_density()+
  geom_freqpoly(aes(y = after_stat(density)))+
  theme_bw()
```


```{r}
actractiveness_by_region
```


```{r}
# %error (using %change/size = ln(a) - ln(b) for small %)
sapply(log(actractiveness_by_region) - estimates , mean)
```

Save the data in R format:
```{r}
r_output = lapply(output, function(jx){
  np$array(jx)
})

saveRDS(r_output, file = "sample_model_1.rds")
```




```{r}
rnorm(100, 0, 4) |> exp() |> 
  ggplot(aes(x))
```




