---
title: "simple_nuts"
output: html_document
---

Let's try the brms package:
```{r}
library(brms)
library(dplyr)
bind <- function(...) cbind(...)

N <- 20

df <- data.frame(
  y1 = rbinom(N, 10, 0.5), y2 = rbinom(N, 10, 0.7), 
  y3 = rbinom(N, 10, 0.9), x = rnorm(N)
) %>%
  mutate(
    size = y1 + y2 + y3,
    y1 = y1 / size,
    y2 = y2 / size,
    y3 = y3 / size
  )
df$y <- with(df, cbind(y1, y2, y3))

df$region = factor(rep(1:4, 5))


bformula = bf(y1 ~ x + (1|region), y2 ~ x)
make_stancode(bf(bind(y1, y2, y3) ~ x + (1|region)), df, dirichlet())
make_standata(bind(y1, y2, y3) ~ x + (1 | region), df, dirichlet())

fit <- brm(bformula, df, dirichlet())
summary(fit)
```




```{r}
library(reticulate)
py_require(packages = c("numpyro",
                        "graphviz",
                        "blackjax",
                        "IPython"))
```


Generate the data - this is the model of electoral data proposed in the paper ...
```{r}
library(MCMCpack)
library(ggplot2)
# library(reshape2)

# Set alpha parameters
alpha <- c(5, 5, 5)

# Generate 500 samples
samples <- rdirichlet(500, alpha)
```

Let's simulate data for a region.
```{r}
# Burgland region
set.seed(123)
# real actractiveness params for the region (alphas)
# 4 parties model
act_burg <- c(0.4, 1, 3, 4)

# burgland has 700 electoral districts
n_electoral <- 700

# get the electoral results
bur_prop <- as.data.frame(rdirichlet(n_electoral, act_burg))
party_names <- c("alpha_party", "beta_party", "gamma_party", "others")
colnames(bur_prop) <- party_names
```


# let's plot parties marginal performance.
```{r}
# for party 1 and 3
bur_prop |> 
  ggplot(aes(x = alpha_party))+
  geom_density()+
  theme_bw()
```


```{r}
# for party 1 and 3
bur_prop |> 
  ggplot(aes(x = gamma_party))+
  geom_density()+
  theme_bw()
```

# Let's build a toy dataset for the entire nation

Simulate 20 different regions with 20 different sizes (# electoral districts).
```{r}
n_regions <- 10

n_district_by_region <- rbinom(n_regions, 500, 0.3)
region_id <- factor(1:n_regions)

# party_national_performance (kinda)
national_alpha = c(0.4, 1, 3, 4)
alpha_by_region <- lapply(national_alpha, function(alpha){
  as.numeric(alpha + pmax(rnorm(n_regions, sd = 2), 0))
})

actractiveness_by_region <- as.data.frame(alpha_by_region)
colnames(actractiveness_by_region) <- party_names
```

Now for each region sample the electoral district vote share distribution:
```{r}
data = data.frame()
for (i in 1:n_regions){
  n = n_district_by_region[i]
  alpha = as.numeric(actractiveness_by_region[i,] )
  print(c(n, alpha))
  vote_share = rdirichlet(n, alpha) 
  vote_share <- as.data.frame(vote_share)
  colnames(vote_share) <- party_names
  vote_share$region = region_id[i]
  data <- rbind(data,vote_share)
}
```

plot some of the vote share distribution
```{r}
data |> 
  ggplot(aes(x = alpha_party))+
  geom_density(fill = "#999999")+
  facet_wrap(.~region)+
  theme_bw()
```

```{r}
data |> 
  ggplot(aes(x = gamma_party))+
  geom_density(fill = "#999999")+
  facet_wrap(.~region)+
  theme_bw()
```


# Now let's find a way to estimate the data by region.
```{python}
from jax import random, vmap
import jax.numpy as jnp
from jax.scipy.special import logsumexp

import numpyro
from numpyro import handlers
from numpyro.diagnostics import hpdi
import numpyro.distributions as dist
from numpyro.infer import MCMC, NUTS

import numpy as np


def model_region(obs = None, n_district = None):
    
    with numpyro.plate("party", 4):
      # log_a = numpyro.sample("log_a", dist.Normal(0, 3))
      a = numpyro.sample("a", dist.HalfNormal(3))
      
    # a = jnp.exp(log_a)
    
    with numpyro.plate("data", 160):
      numpyro.sample("obs", dist.Dirichlet(a), obs = obs)
        

```

```{r}
nuts_kernel = py$NUTS(py$model_region)

mcmc = py$MCMC(nuts_kernel, num_samples=2000L, num_warmup=2000L, jit_model_args = TRUE )
rng_key = py$random$key(42L)

# Get the data in jax arrays
data_jax = py$jnp$array(as.matrix(data[data$region == 3, 1:4]))
data_jax$shape[1] |> print()

mcmc$run(rng_key, obs = data_jax, n_district = py$jnp$array(160L))
```

```{r}
mcmc$print_summary()

c(-0.72, 0.03, 1.16, 1.92) |> exp()
c(0.08, 0.06, 1.65, 1.37) |> exp()
c(-0.97, -0.03, 1.73, 1.48) |> exp()
```


# Now lejt's build the hierarchical model.
```{python}
# def model_national(obs = None, region_idx = None, N = None):
#     
#     # Make the prior for the population dirichlet
#     # with numpyro.plate("party", 4):
#     #   a = numpyro.sample("a", dist.HalfNormal(1))
#       
#     with numpyro.plate("region", len(np.unique(region_idx))):
#       a_reg = numpyro.sample("a_reg", dist.Dirichlet)
#       
#     with numpyro.plate("data", N):
#       numpyro.sample("obs", dist.Dirichlet(a_reg[region_idx]), obs = obs)
#       
      
def model_national(obs = None, region_idx = None, N = None):
    
    sigma_reg = numpyro.sample("sigma_reg", dist.HalfNormal(1))
    
    with numpyro.plate("party", 4):
      log_c_party = numpyro.sample("c_party", dist.Normal(jnp.zeros(1), jnp.ones(1)*3))
    
    c_party = jnp.exp(log_c_party)
    
    with numpyro.plate("party", 4):
      with numpyro.plate("region", len(np.unique(region_idx))):
        log_a_reg = numpyro.sample("a_reg", dist.Normal(c_party, sigma_reg))
      
    a_reg = jnp.exp(log_a_reg)
    
    with numpyro.plate("data", N):
      numpyro.sample("obs", dist.Dirichlet(a_reg[region_idx]), obs = obs)
```

```{r}
nuts_kernel = py$NUTS(py$model_national)

mcmc = py$MCMC(nuts_kernel, num_samples=2000L, num_warmup=1000L, jit_model_args = TRUE)
rng_key = py$random$key(42L)

# Get the data in jax arrays
data_jax = py$jnp$array(as.matrix(data[, 1:4]))
region_idx =  py$jnp$array(as.integer(data[, 5]))
```


# Move it to blackjax
```{python}
from numpyro.infer.util import initialize_model

rng_key, init_key = random.split(r.rng_key)

init_params, potential_fn_gen, *_ = initialize_model(init_key, model_national, model_args=(r.data_jax, r.region_idx, 1526), dynamic_args=True, )
```

Convert it to a logdensity function:
```{python}
from jax import jit

logdensity_fn = lambda position: -potential_fn_gen(r.data_jax, r.region_idx, 1526)(position)
logdensity_jit = jit(logdensity_fn)
initial_position = init_params.z
```

Create the kernel for nuts algo with step size and inv mass matrix
```{python}
import blackjax

inv_mass_matrix = jnp.ones(45)
step_size = 1e-3

nuts = blackjax.nuts(logdensity_jit, step_size, inv_mass_matrix)
init_state = nuts.init(initial_position)
kernel = nuts.step

```
 the inference loop
```{python}
import jax

def inference_loop(rng_key, kernel, initial_state, num_samples):
  @jax.jit
  def one_step(state, rng_key):
      state, _ = kernel(rng_key, state)
      return state, state

  keys = jax.random.split(rng_key, num_samples)
  _, states = jax.lax.scan(one_step, initial_state, keys)

  return states
```

Run the algo:
```{python}
n_samples = 100
rng_key, sample_key = random.split(rng_key)
states = inference_loop(sample_key, kernel, init_state, n_samples)

mcmc_samples = states.position
```

Save the results of the test to file:
```{python}
# 
# import pickle
# import os
# 
# path = "C:/Users/Ermenegisto/CODING/R/UNI/CSS/final_proj/dirichlet_estimation"
# with open(path, 'wb') as file:
#     pickle.dump(mcmc_samples, file)
```



# Let's look at the results in r
```{r}
np = import("numpy", as = "np")
output = py$mcmc_samples

# population level estimates, prior
data_pop = as.data.frame(np$array(output$c_party))
names(data_pop) <- c("a1","a2","a3","a4")

library(tidyr)

long <- data_pop |> # [1000:nrow(data_pop)] %>%
  pivot_longer(
    cols = a1:a4, 
    names_to = "alpha",
    values_to = "value"
)

alpha
exp(sapply(data_pop, mean))

long |> 
  ggplot(aes(x = exp(value)))+
  geom_density()+
  facet_wrap(.~alpha)+
  theme_bw()
```

And now let's plot the region based ones
```{r}
data_reg = output$a_reg

n_samples = data_reg$shape[[1]]
burnout = 10

data_plot = data_reg[burnout:n_samples,,]
# mean_concentration per region x party
estimates = as.data.frame(np$array(data_plot$mean(0L)))
names(estimates) <- party_names
exp(estimates)
```


```{r}
actractiveness_by_region
```


```{r}
# %error (using %change/size = ln(a) - ln(b) for small %)
sapply(log(actractiveness_by_region) - estimates , mean)
```

Save the data in R format:
```{r}
r_output = lapply(output, function(jx){
  np$array(jx)
})

saveRDS(r_output, file = "sample_model_1.rds")
```




```{r}
rnorm(100, 0, 4) |> exp() |> 
  ggplot(aes(x))
```




