---
title: "simple_nuts"
output: html_document
---

```{r}
library(reticulate)
py_require(packages = c("numpyro"))
```


Generate the data - this is the model of electoral data proposed in the paper ...
```{r}
library(MCMCpack)
library(ggplot2)
# library(reshape2)

# Set alpha parameters
alpha <- c(5, 5, 5)

# Generate 500 samples
samples <- rdirichlet(500, alpha)
```

Let's simulate data for a region.
```{r}
# Burgland region
set.seed(123)
# real actractiveness params for the region (alphas)
# 4 parties model
act_burg <- c(0.4, 1, 3, 4)

# burgland has 700 electoral districts
n_electoral <- 700

# get the electoral results
bur_prop <- as.data.frame(rdirichlet(n_electoral, act_burg))
party_names <- c("alpha_party", "beta_party", "gamma_party", "others")
colnames(bur_prop) <- party_names
```


# let's plot parties marginal performance.
```{r}
# for party 1 and 3
bur_prop |> 
  ggplot(aes(x = alpha_party))+
  geom_density()+
  theme_bw()
```


```{r}
# for party 1 and 3
bur_prop |> 
  ggplot(aes(x = gamma_party))+
  geom_density()+
  theme_bw()
```

# Let's build a toy dataset for the entire nation

Simulate 20 different regions with 20 different sizes (# electoral districts).
```{r}
n_regions <- 10

n_district_by_region <- rbinom(n_regions, 500, 0.3)
region_id <- factor(1:n_regions)

# party_national_performance (kinda)
national_alpha = c(0.4, 1, 3, 4)
alpha_by_region <- lapply(national_alpha, function(alpha){
  as.numeric(alpha + pmax(rnorm(n_regions, sd = 2), 0))
})

actractiveness_by_region <- as.data.frame(alpha_by_region)
colnames(actractiveness_by_region) <- party_names
```

Now for each region sample the electoral district vote share distribution:
```{r}
data = data.frame()
for (i in 1:n_regions){
  n = n_district_by_region[i]
  alpha = as.numeric(actractiveness_by_region[i,] )
  print(c(n, alpha))
  vote_share = rdirichlet(n, alpha) 
  vote_share <- as.data.frame(vote_share)
  colnames(vote_share) <- party_names
  vote_share$region = region_id[i]
  data <- rbind(data,vote_share)
}
```

plot some of the vote share distribution
```{r}
data |> 
  ggplot(aes(x = alpha_party))+
  geom_density()+
  facet_wrap(.~region)+
  theme_bw()
```

```{r}
data |> 
  ggplot(aes(x = gamma_party))+
  geom_density()+
  facet_wrap(.~region)+
  theme_bw()
```


# Now let's find a way to estimate the data by region.








